Deep Learning

### 2015-01-31 deepreview.tex
higher-resolution figures, removed old bp figure, rewrote preface

From the earliest days of pattern recognition~\citep{selfridge,Rosenblatt57}
researchers aimed to replace hand-engineered features by multiple layers of
learned features.  This was eventually achieved by using the
backpropagation algorithm which was discovered independently by several
different groups in the 1970s and 1980s~\citep{Werbos74,Parker85,LeCun85,RHW}.  
Backpropagation is simply an application of
the chain rule of derivatives (left of Figure~\ref{fig:backprop-box}),
which yields an efficient way of computing the gradient of a
smooth objective function with respect to all of the weights 
%% and parameters 
%% FOR THE NOVICE READER INCLUDING BOTH WEIGHTS AND "PARAMTERS" IS CONFUSING.
%% ITS BETTER TO SAY SOMETHING CLEARER EVEN IF IT IS LESS GENERAL OR CORRECT.
in  a layered network (Figures~\ref{fig:backprop-box} and~\ref{fig:mlp}). 
The objective function typically measures an error to be minimized: 
the discrepancy between
the actual output of the network on a given training case and the target
output for that case.  The key insight is that the error derivatives with
respect to the {\it ouputs} of a layer can be computed by working
backwards from the error derivatives with respect to its {\em inputs}
to the subsequent layer.
%% THE TRICKY STEP IS COMPUTING dE/dOUT FROM dE/dIN IN THE SUBSEQUENT
%% LAYER.  I THINK WE SHOULD FOCUS ON GETTING OVER THAT IDEA FIRST.
%%  I PLAN TO REDO THE FIGURE TO MAKE IT SIMPLER.  ITS MUCH TOO
%%  CONFUSING FOR A NOVICE AT PRESENT.  WE SHOULD HAVE A SEPARATE
%%  DISCUSSION OF THAT.
Once this backwards recursion has been used to compute
error derivatives for the outputs of all layers, it is easy to compute
error derivatives for the incoming weights of the neurons in each layer 
(Figure~\ref{fig:mlp}).

@@

When backpropagation was first introduced, its most exciting use was for
training recurrent neural networks. These receive input vectors at each
time-step and have one or more layers of hidden units that, in addition to
having connections to higher layers or ouput units, also have connections
that allow the neurons within a layer to receive weighted input from those
same neurons at the previous time-step.  The gradient with respect to a recurrent
weight, $w_{ij}$ that allows neuron $i$ at time $t$ to influence the
activity of neuron $j$ at time $t+1$ can be computed by
backpropagation-through-time (BPTT) of the error derivative for $j$ at time
$t+1$. The full gradient with respect to $w_{ij}$ on a whole input sequence is just the
sum of its gradients at each time-step.

@@

\begin{figure}[ht]
\fbox{
\begin{minipage}{\textwidth}
\centerline{
\includegraphics[width=0.49\linewidth]{bp-mlp1.pdf}
\includegraphics[width=0.49\linewidth]{bp-mlp5.pdf}
}
\centerline{
\includegraphics[width=0.89\linewidth]{fig-hidden-recurrence-rnn.pdf}
}
\caption{Top left: graph of feedforward computations in a multi-layer
neural network with one input layer ($x$), two hidden layers ($h_1$ and $h_2$),
one output layer (out), a target $y$, and an error that compares the output
and the target. The synaptic weight matrices $W_1$, $W_2$ and $W_3$ are
the parameters of the neural network.\newline
Top right: back-propagation through the same multi-layer neural network
(downward arrows compute the partial derivative of the error with respect
to every node, i.e., the units value and the parameters).\newline
Bottom: a recurrent neural network (left) and the unfolding in time of the
computation involved in its forward computation. The artificial neurons
(e.g. hidden units grouped under node $s$ with values $s_t$ at time $t$) 
get inputs from other neurons at previous time steps
(this is represented with the black square, standing for a delay of
one time step, on the left). In this way, a recurrent net can map an input sequence
with elements $x_t$ into an output sequence with elements $o_t$, with
each $o_t$ depending on all the previous $x_{t'}$ (for $t'\leq t$). The
same parameters (matrices $U$, $V$, $W$) are used at each time step.
Many other architectures are possible, including a variant in which
the sampled outputs (e.g. words) can be generated according to an output 
distribution (the probability of the next word, for every possible next word) 
and used as inputs for the next time step.
The back-propagation algorithm (Figure~\ref{fig:backprop-box})
can be directly applied to the computational graph of the unfolded network
on the right hand side, to compute the derivative of a total error
(e.g., the log-probability of generating the right sequence of
words) with respect to all the states $s_t$ and all the parameters.
}
\label{fig:mlp}
\end{minipage}
}
\end{figure}

>>>>

From the earliest days of pattern recognition~\citep{selfridge,Rosenblatt57}
researchers aimed to replace hand-engineered features by multiple layers of
learned features.  This was eventually achieved by using the
backpropagation algorithm which was discovered independently by several
different groups in the 1970s and 1980s~\citep{Werbos74,Parker85,LeCun85,RHW}.  
Backpropagation is simply an application of
the chain rule of derivatives (left of Figure~\ref{fig:backprop-box}),
which yields an efficient way of computing the gradient of a
smooth objective function with respect to all of the weights 
%% and parameters 
%% FOR THE NOVICE READER INCLUDING BOTH WEIGHTS AND "PARAMTERS" IS CONFUSING.
%% ITS BETTER TO SAY SOMETHING CLEARER EVEN IF IT IS LESS GENERAL OR CORRECT.
in  a layered network (Figure~\ref{fig:backprop-box}). 
The objective function typically measures an error to be minimized: 
the discrepancy between
the actual output of the network on a given training case and the target
output for that case.  The key insight is that the error derivatives with
respect to the {\it ouputs} of a layer can be computed by working
backwards from the error derivatives with respect to its {\em inputs}
to the subsequent layer.
%% THE TRICKY STEP IS COMPUTING dE/dOUT FROM dE/dIN IN THE SUBSEQUENT
%% LAYER.  I THINK WE SHOULD FOCUS ON GETTING OVER THAT IDEA FIRST.
%%  I PLAN TO REDO THE FIGURE TO MAKE IT SIMPLER.  ITS MUCH TOO
%%  CONFUSING FOR A NOVICE AT PRESENT.  WE SHOULD HAVE A SEPARATE
%%  DISCUSSION OF THAT.
Once this backwards recursion has been used to compute
error derivatives for the outputs of all layers, it is easy to compute
error derivatives for the incoming weights of the neurons in each layer 
(Figure~\ref{fig:backprop-box}).

@@

When backpropagation was first introduced, its most exciting use was for
training recurrent neural networks. These receive input vectors at each
time-step and have one or more layers of hidden units that, in addition to
having connections to higher layers or ouput units, also have connections
that allow the neurons within a layer to receive weighted input from those
same neurons at the previous time-step.  As illustrated in
Figure~\ref{fig:rnn}, the gradient with respect to a recurrent
weight, $w_{ij}$ that allows neuron $i$ at time $t$ to influence the
activity of neuron $j$ at time $t+1$ can be computed by
backpropagation-through-time (BPTT) of the error derivative for $j$ at time
$t+1$. The full gradient with respect to $w_{ij}$ on a whole input sequence is just the
sum of its gradients at each time-step.

@@

\begin{figure}[ht]
\fbox{
\begin{minipage}{\textwidth}
%\centerline{
%\includegraphics[width=0.49\linewidth]{bp-mlp1.pdf}
%\includegraphics[width=0.49\linewidth]{bp-mlp5.pdf}
%}
\centerline{
\includegraphics[width=0.89\linewidth]{fig-hidden-recurrence-rnn.pdf}
}
\caption{
%Top left: graph of feedforward computations in a multi-layer
%neural network with one input layer ($x$), two hidden layers ($h_1$ and $h_2$),
%one output layer (out), a target $y$, and an error that compares the output
%and the target. The synaptic weight matrices $W_1$, $W_2$ and $W_3$ are
%the parameters of the neural network.\newline
%Top right: back-propagation through the same multi-layer neural network
%(downward arrows compute the partial derivative of the error with respect
%to every node, i.e., the units value and the parameters).\newline
A recurrent neural network (left) and the unfolding in time of the
computation involved in its forward computation. The artificial neurons
(e.g. hidden units grouped under node $s$ with values $s_t$ at time $t$) 
get inputs from other neurons at previous time steps
(this is represented with the black square, standing for a delay of
one time step, on the left). In this way, a recurrent net can map an input sequence
with elements $x_t$ into an output sequence with elements $o_t$, with
each $o_t$ depending on all the previous $x_{t'}$ (for $t'\leq t$). The
same parameters (matrices $U$, $V$, $W$) are used at each time step.
Many other architectures are possible, including a variant in which
the sampled outputs (e.g. words) can be generated according to an output 
distribution (the probability of the next word, for every possible next word) 
and used as inputs for the next time step.
The back-propagation algorithm (Figure~\ref{fig:backprop-box})
can be directly applied to the computational graph of the unfolded network
on the right hand side, to compute the derivative of a total error
(e.g., the log-probability of generating the right sequence of
words) with respect to all the states $s_t$ and all the parameters.
}
\label{fig:rnn}
\end{minipage}
}
\end{figure}
